A Systematic Literature Review on Extensions to Role Based Access Control

### 2013-04-25 paper/sections/analysis.tex
Minor formatting clean-up of analysis section.

The RBAC standard model provides an abstraction of authorization based on user role assignment (UA) and role
permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
security mechanism, a variety of extended RBAC models have been proposed over the years to meet
security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.

@@

@@

The RBAC standard was designed with enterprises in mind such that when practioners implemented RBAC into their systems
there would be a reasonable assurance being based off a well thought out model.  As extensions to the standard model
come along, thought and time should be given to how features and nuances of their models may impact implementation
in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as 4 papers did, should be
a prototype implementation of the model for review by both pracitioners and researchers. Of the models that produced an 
implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.

@@

The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
evaluation types and which primary sources provided which type.  In some cases a single primary source provided mutliple evaluation
types.

@@

When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
Further, in the case of extensions to the RBAC standard model, the model should be accompanied by validation of the model as a stand-alone
access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
extenion models are lacking.

The primary source of validation a developer or practioner may encounter is a qualitative discussion of real-world
scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rarer cases, 
one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
of how an extension model handles a real-world scenario provides developers and practioners ancedotal evidence at best
for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
model, developers may be left implementing a more complex model to address their requirements when the standard model would
have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
case studies of a model in action provide developers with evidence that the model works as intended when applied.

@@

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reasaon for the explosion of
RBAC extensions.  The authors identified domains presented within the primary sources by looking for example
scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
We found that the domains mentioned and their associated sources are:

@@

Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
was introduced. 
The domain analysis results produced 5 papers that targeted extensions that are designed to address the requirements of mobile computing.
For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.

>>>>

@@

The RBAC standard model provides an abstraction on top of authorization based on user role assignment (UA) and role
permission assignment (PA). Since this simple model may not provide a fine-grained access control for sophisticated
security mechanism, a variety of extended RBAC models have been proposed over the years to meet
security requirements. When administrators design a model, it is important to capture an important abstraction to help the model to be enforced in a system.

@@

The RBAC standard was designed with enterprises in mind such that when practitioners implemented RBAC into their systems
there would be a reasonable assurance being based off a well thought out model.  As extensions to the standard model
come along, thought and time should be given to how features and nuances of their models may impact implementation
in order to achieve the same goals as the original standard.  The primary sources should a significant lack of implementation
with over 70\% of the models having no notion of attempting to implement them.  The bare minimum, as 4 papers did, should be
a prototype implementation of the model for review by both practitioners and researchers. Of the models that produced an 
implementation within the enterprise world, two were from within the medical domain and two were implemented using web application technologies.

@@

The 28 primary sources were examined for evidence that evaluations of the proposed model were presented by the model authors.
Further, we identified a set of evaluation types found within the each of the primary sources and provide below a list of the 
evaluation types and which primary sources provided which type.  In some cases a single primary source provided multiple evaluation
types.

@@

When proposing an access control model, providing an evaluation of the model is a key component in establishing the validity of the model. 
Further, in the case of extensions to the RBAC standard model, the model should be accompanied by validation of the model as a stand-alone
access control model and in comparison to the model upon which the enhancements are being made. The results show that robust evaluations of
extension models are lacking.

The primary source of validation a developer or practitioner may encounter is a qualitative discussion of real-world
scenarios and how the proposed model can tackle those situations. In rare cases, five primary sources, the model 
authors provide some discussion of how the RBAC standard model is deficient in tackling the scenario. In rarer cases, 
one paper from our results, a case study is performed to examine how the proposed model works in practice. Discussions
of how an extension model handles a real-world scenario provides developers and practitioners anecdotal evidence at best
for what types of situations the proposed model could handle. Further, by not providing a comparison to the RBAC standard
model, developers may be left implementing a more complex model to address their requirements when the standard model would
have sufficed.  Further, given the nature of access control models as grounded in application to enterprise implementations, 
case studies of a model in action provide developers with evidence that the model works as intended when applied.

@@

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
RBAC extensions.  The authors identified domains presented within the primary sources by looking for example
scenarios cast within a particular domain or mention of domain requirements within the body of the paper.
We found that the domains mentioned and their associated sources are:

@@

Mobile computing has seen a dramatic increase in the number of available devices, operating systems and applications since 1997 when the first smart phone
was introduced. 
The domain analysis results produced five papers that targeted extensions that are designed to address the requirements of mobile computing.
For a domain that has roots in personal and enterprise computing, protecting the data of both through access controls is paramount given their ubiquity.
